GDPR: Is non-compliance profitable for Meta Platforms, Inc.?

### 2023-05-23 docs/response-time-research.md

# Accelerating response times for high-traffic social media platforms

By Kostadinov, Stoyan S.L.

<!-- TOC -->
* [Accelerating response times for high-traffic social media platforms](#accelerating-response-times-for-high-traffic-social-media-platforms)
  * [Introduction](#introduction)
    * [Context](#context)
    * [Research questions and objectives](#research-questions-and-objectives)
    * [Methodology](#methodology)
    * [Constraints](#constraints)
    * [Glossary](#glossary)
  * [Literature review](#literature-review)
  * [Non-functional testing](#non-functional-testing)
  * [References](#references)
<!-- TOC -->

## Introduction

### Context

The majority of content on Reddit consists of text-based and media-based (videos and images) posts.
According to recent statistics, Reddit received an average of 30 billion monthly views in 2020,
making it the 6th largest social media platform in terms of
traffic.[[x]](https://www.businessofapps.com/data/reddit-statistics/)[[x]](https://www.similarweb.com/top-websites/computers-electronics-and-technology/social-networks-and-online-communities/)
As Reddit's revenue model heavily relies on
advertising[[x]](https://www.businessofapps.com/data/reddit-statistics/), the company stands to earn
more money with increased user visits to the platform.

Website speed is widely recognized as a crucial factor in user experience, with even a 1-second
delay in page load time leading to a 7% decrease in
conversions.[[x]](https://www.truconversion.com/traffic/decrease-page-load-time/)[[x]](https://cxl.com/blog/11-low-hanging-fruits-for-increasing-website-speed-and-conversions/)
When considering Reddit's annual revenue of $350
million[[x]](https://www.businessofapps.com/data/reddit-statistics/), this conversion loss
translates to a potential revenue loss of $24.5 million per year.

These factors clearly indicate that Reddit and other social media platforms have a strong incentive
to optimize their response times.

### Research questions and objectives

The goal of this research paper is to gain insight into different optimization techniques and
provide recommendations for the _not-reddit_[[x]](https://github.com/stoyanK7/not-reddit/) social
media platform.

> _not-reddit_ is a social news aggregation platform with sub-communities, messaging, and awards
> that allows users to post and share content, vote on submissions, and comment on posts.

---

> _**How can response times for the not-reddit social media platform be accelerated?**_

In order go get a better understanding of the problem, the following sub-questions will be
addressed:

> - _How does a request at not-reddit currently look like?_
> - _How does a request at not-reddit currently perform?_
> - _What performance-improving techniques do major social media platforms use?_
> - _How can the impact different techniques' be measured and evaluated?_

### Methodology

To achieve the research objectives and address the research questions, the following methodology
will be employed:

1. **Literature study** - A literature review will be conducted to gain insight into the
   performance-improving techniques used by major social media platforms.
2. **Non-functional testing** - A series of non-functional tests will be conducted to measure the
   performance of the not-reddit platform before and after applying the performance-improving
   techniques.
3. **IT architecture sketching** - A sketch of the not-reddit IT architecture will be created to
   identify the components that can be optimized.

> Note: The above methods are taken from ICT Research
> methods.[[x]](https://ictresearchmethods.nl/Methods)

### Constraints

The research is constrained by the following factors:

- **Time**\
  The research is limited to 3 weeks.
- **Resources**\
  The research is limited to the resources available to the author. All tests are ran from
  a `ASUSTeK COMPUTER INC. ZenBook UX434FLC_UX434FL` laptop with `8GB` of RAM and
  an `Intel(R) Core(TM) i7-10510U CPU @ 1.80GHz` processor.
- **Scope**\
  The research is limited to the not-reddit platform and does not consider other social media
  platforms.
- **Data**\
  The research is limited to the data available to the author. The not-reddit platform is a
  prototype and does not have any real users. The data used for the research is generated by
  the author.
- **Expertise**\
  The research is limited to the author's expertise. The author is a student with limited
  experience in the field of performance optimization.

### Glossary

| Term | Definition |
|------|------------|
| xxx  |            |

## Baseline

![img](img/2023-05-22-architecture-container.png)

## Literature review

## Non-functional testing

## References

TODO

### 2023-05-23 docs/response-time-research.md

# Accelerating response times for high-traffic social media platforms

By Kostadinov, Stoyan S.L.

<!-- TOC -->

* [Accelerating response times for high-traffic social media platforms](#accelerating-response-times-for-high-traffic-social-media-platforms)
    * [Introduction](#introduction)
        * [Context](#context)
        * [Research questions and objectives](#research-questions-and-objectives)
        * [Methodology](#methodology)
        * [Constraints](#constraints)
        * [Glossary](#glossary)
    * [Literature review](#literature-review)
    * [Non-functional testing](#non-functional-testing)
    * [References](#references)

<!-- TOC -->

@@

### Glossary

| Term | Definition |
|------|------------|
| xxx  |            |

## Baseline

On a system level, the not-reddit platform consists of 4 components - not-reddit system itself,
Azure Active Directory (AAD), Azure Blob Storage (ABS), and Azure Email Communication Service (
AECS).
![img](img/2023-05-22-architecture-system.png)

The scope of this research is focuses on the post service which is responsible for handling creation
and retrieval of posts and media. A typical request starts from the client (browser), goes
through the API gateway and lands in the Post API. The API will retrieve the post from the database
and send it back to the client. Depending of the `type` of the post, the client will send additional
request for the media. In total, for a media post, 2 requests will be made.
![img](img/2023-05-23-architecture-post.png)

### 2023-05-23 docs/response-time-research.md

# Accelerating response times for high-traffic social media platforms

By Kostadinov, Stoyan S.L.

<!-- TOC -->

* [Accelerating response times for high-traffic social media platforms](#accelerating-response-times-for-high-traffic-social-media-platforms)
    * [Introduction](#introduction)
        * [Context](#context)
        * [Research questions and objectives](#research-questions-and-objectives)
        * [Methodology](#methodology)
        * [Constraints](#constraints)
        * [Glossary](#glossary)
    * [Baseline](#baseline)
        * [Baseline results](#baseline-results)
    * [Literature review](#literature-review)
    * [Non-functional testing](#non-functional-testing)
    * [References](#references)

@@

> Note: The above methods are taken from ICT Research
> methods.[[x]](https://ictresearchmethods.nl/Methods)

### Constraints

The research is constrained by the following factors:

- **Time**\
  The research is limited to 3 weeks.
- **Resources**\
  The research is limited to the resources available to the author. All tests are ran from
  a `ASUSTeK COMPUTER INC. ZenBook UX434FLC_UX434FL` laptop with `8GB` of RAM and
  an `Intel(R) Core(TM) i7-10510U CPU @ 1.80GHz` processor. The Azure Kubernetes Services cluster is
  running on a `Standard_B4ms` machine with 1 node pool.
- **Scope**\
  The research is limited to the not-reddit platform and does not consider other social media
  platforms.
- **Data**\
  The research is limited to the data available to the author. The not-reddit platform is a
  prototype and does not have any real users. The data used for the research is generated by
  the author.
- **Expertise**\
  The research is limited to the author's expertise. The author is a student with limited
  experience in the field of performance optimization.

The landing page will contain 6 text and 4 media posts.

### Glossary
TODO: look for terms at the end of the document and add them here

| Term                            | Definition                                                                                                                                                                 |
|---------------------------------|----------------------------------------------------------------------------------------------------------------------------------------------------------------------------|
| Azure Kubernetes Services (AKS) | A fully managed container orchestration service provided by Microsoft Azure, enabling developers to deploy, scale, and manage containerized applications using Kubernetes. |

## Measurements

The following measurements will be used to evaluate the performance of the not-reddit platform:

- **Latency**\
  The time it takes for a request to be sent from the client to the server and back.
- **Throughput**\
  The number of requests that can be processed per second.
- **Error rate**\
  The percentage of requests that fail.
- **CPU usage**\
  The percentage of CPU used by the application.
- **Network usage**\
  The percentage of network bandwidth used by the application.

Latency, Throughput and Error rate are measured using _Apache
JMeter_[[x]](https://jmeter.apache.org/) and _Azure Kubernetes
Metrics_.[[x]](https://learn.microsoft.com/en-us/azure/aks/monitor-aks)

@@

## Baseline

On a system level, the not-reddit platform consists of 4 components - not-reddit system itself,
Azure Active Directory (AAD), Azure Blob Storage (ABS), and Azure Email Communication Service (
AECS).
![img](img/2023-05-22-architecture-system.png)
<p align="center"><i>Figure 1: System architecture</i></p>

The scope of this research is focuses on the post service which is responsible for handling creation
and retrieval of posts and media. A typical request starts from the client (browser), goes
through the API gateway and lands in the Post API. The API will retrieve the post from the database
and send it back to the client. Depending of the `type` of the post, the client will send additional
request for the media. In total, for a media post, 2 requests will be made.
![img](img/2023-05-23-architecture-post.png)
<p align="center"><i>Figure 2: Post service architecture</i></p>

### Baseline results

### 2023-05-24 docs/response-time-research.md

The research is constrained by the following factors:

- **Time**\
  The research is limited to 3 weeks.
- **Resources**\
  The research is limited to the resources available to the author. All tests are ran from
  a `ASUSTeK COMPUTER INC. ZenBook UX434FLC_UX434FL` laptop with `8GB` of RAM and
  an `Intel(R) Core(TM) i7-10510U CPU @ 1.80GHz` processor. The Azure Kubernetes Services cluster is
  running on a `Standard_B4ms` machine with 1 node pool.
- **Scope**\
  The research is limited to the not-reddit platform and does not consider other social media
  platforms.
- **Data**\
  The research is limited to the data available to the author. The not-reddit platform is a
  prototype and does not have any real users. The data used for the research is generated by
  the author.
- **Expertise**\
  The research is limited to the author's expertise. The author is a student with limited
  experience in the field of performance optimization.

The landing page will contain 6 text and 4 media posts.

#### Measurements

- **Latency**\
  The time it takes for a request to be sent from the client to the server and back.
- **Throughput**\
  The number of requests that can be processed per second.
- **Error rate**\
  The percentage of requests that fail.

All are measured using _Apache JMeter_.[[x]](https://jmeter.apache.org/) A total of 5 requests
different requests will be made to simulate real user behavior:

- /api/post?sort_by=hot&page=0
- /api/post/media/7.png
- /api/post/media/8.png
- /api/post/media/9.png
- /api/post/media/10.png

The number of threads will be set to 1000 and the ramp-up period to 1 second.

### Glossary

TODO: look for terms at the end of the document and add them here

### 2023-06-09 docs/research/LaTeX/research.tex
Add script for transforming data of gdpr fines
Finish research introduction

\documentclass[12pt, a4paper]{article}
\usepackage{indentfirst}
\usepackage{graphicx}
\graphicspath{{images/}}
\title{GDPR: Is non-compliance profitable for Meta Platforms, Inc.?}
\author{Stoyan L. Kostadinov}
\date{June 2023}
\begin{document}
\maketitle

@@

GDPR, Facebook, Meta, profit, non-compliance, big tech, privacy

@@

In the recent years, the increasing usage of the Internet has lead to a rapid
expansion of data collection and processing\cite{}. As concerns regarding data
privacy and protection grew, the European Union (EU) implemented the General
Data Protection Regulation (GDPR) in May 2018\cite{}.

GDPR aimed to safeguard the privacy rights of individuals within the EU by
regulating the collection, storage, and transfer of personal data\cite{}.
Notably, data should be collected for specified, explicit and legitimate
purposes\cite{} and data can be transferred outside the EU if general data
processing principles are in place \cite{}. Compliance with GDPR has become a
critical issue for big-tech companies, which often handle massive amounts of
user data.

One of those companies is \textit{Meta Platforms, Inc.} (formerly
\textit{Facebook, Inc.}), which has been fined numerous times for GDPR
violations and amounts for more than half of the total currency value of all
GDPR fines\cite{}. The company has also faced legal challenges and public
backlash over its data collection practices\cite{}. This has raised questions
about the financial impact of GDPR non-compliance on \textit{Meta Platforms,
Inc.} and whether it is profitable for the company to continue violating the
regulation.

@@

Most of the numerical data is coming from the Meta Investor Relations
website\cite{}.

This research is also following the protocol of Prisma flowchart for systematic
reviews\cite{}. Google Scholar was used to search for articles containing the
keywords above. The first 360 results were considered. The articles were
filtered by relevance following their titles and 60 were considered. The
abstracts of the 60 articles were read and X were considered
% TODO: Describe flowchart with sentences above here TODO: Put numbers and add
% flowchart here as Figure 1

@@

\begin{itemize}
    \item Articles published in English
    \item Articles published between 2018 and 2023
    \item Articles that discuss GDPR
    \item Articles that discuss Meta Platforms, Inc.
    \item Articles that discuss the profit of Meta Platforms, Inc.
    \item Articles that discuss the non-compliance of Meta Platforms, Inc.
\end{itemize}

@@

\begin{itemize}
    \item Articles published in languages other than English
    \item Articles published before 2018
    \item Articles that do not discuss GDPR
    \item Articles that do not discuss Meta Platforms, Inc.
    \item Articles that discuss about ethics
    \item Articles that discuss about the impact of GDPR
    \item Articles that discuss about the impact of Meta Platforms, Inc.
\end{itemize}

>>>>

\documentclass[12pt, a4paper]{article}
\usepackage{indentfirst}
\usepackage{graphicx}
\usepackage{url}
\graphicspath{{images/}}
\title{GDPR: Is non-compliance profitable for Meta Platforms, Inc.?}
\author{Stoyan L. Kostadinov}
\date{June 2023}
\begin{document}
\maketitle

@@

GDPR, Facebook, Meta, profit, non-compliance, big tech, privacy, financial
impact

@@

In the recent years, the increasing usage of the Internet has lead to a rapid
expansion of data collection and processing\cite{khan2014big}. As concerns
regarding data privacy and protection grew, the European Union (EU) implemented
the General Data Protection Regulation (GDPR) in May
2018\cite{greengard2018weighing, historyGdpr}.

GDPR aimed to safeguard the privacy rights of individuals within the EU by
regulating the collection, storage, and transfer of personal
data\cite{historyGdpr}. Notably, data should be collected for specified,
explicit and legitimate purposes\cite{europeanParliamentGdprArticle5} and data
can be transferred outside the EU if general data processing principles are in
place \cite{europeanParliamentGdprArticle44}. Compliance with GDPR has become a
critical issue for big-tech companies, which often handle massive amounts of
user data.

One of those companies is \textit{Meta Platforms, Inc.} (formerly
\textit{Facebook, Inc.}), which has been fined numerous times for GDPR
violations\cite{mrevzar2023analysis}. This has raised questions about the
financial impact of GDPR non-compliance from \textit{Meta} and whether it is
profitable for the company to continue violating the regulation.

The goal of this paper is to investigate the financial impact of GDPR
non-compliance on \textit{Meta Platforms, Inc.} and to determine whether it is
profitable for the company to continue violating the regulations.

@@

All numerical data about \textit{Meta Platforms, Inc.} has been gathered from
the official \textit{Meta Investor Relations} website\cite{fbMetaFinancials}.

GDPR fines imposed on \textit{Meta Platforms, Inc.} have been gathered from the
\textit{GDPR Enforcement Tracker}
website\cite{enforcementtrackerGDPREnforcement}.

This research is following the protocol of \textit{Prisma flowchart} for
systematic reviews\cite{prismaFlowchart}.

\textit{Google Scholar} was used to search for articles containing the keywords
above. The first 360 results were considered. The articles were filtered by
relevance following their titles and 60 were considered. The abstracts of the 60
articles were read and X were considered
% TODO: Describe flowchart with sentences above here TODO: Put numbers and add
% flowchart here as Figure 1

@@

\begin{itemize}
    \item Articles published in English
    \item Articles published between 2018 and 2023
    \item Articles that discuss GDPR
    \item Articles that discuss Meta Platforms, Inc.
    \item Data between 2021 and 2023
\end{itemize}

@@

\begin{itemize}
    \item Articles published in languages other than English
    \item Articles published before 2018
    \item Articles that do not discuss GDPR
    \item Articles that do not discuss Meta Platforms, Inc.
    \item Articles that discuss about ethics
    \item Data before 2021
\end{itemize}
